---
title: "main"
output: html_document
date: "2024-03-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ExData_Plotting1
This is the main R markdown file for the Exploratory Data Analysis course project 1. The goal of this project is to examine how household energy usage varies over a 2-day period in February, 2007. The data is taken from the UCI Machine Learning Repository and is available at the following link: https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip

The individual plots are created using functions defined in the `plot1.R`, `plot2.R`, `plot3.R`, and `plot4.R` files. The data is processed in the `main.R` file.

## Introduction
First, we will download and unzip the data. Then, we will load the data and subset it to the relevant time period. Finally, we will create the plots.

## Data Processing
```{r}
# Download and unzip the data
data_source <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip"
zip_file <- "./data/household_power_consumption.zip"
file_path <- "./data/household_power_consumption.txt"

# If data directory does not exist, create it
if (!file.exists("./data")) {
    dir.create("./data")
}

# If the file does not exist, download it and unzip it
if (!file.exists(file_path)) {
    download.file(data_source, zip_file, method = "curl")
    unzip(zip_file, exdir = "./data")
}

# Load the data
data <- read.table("./data/household_power_consumption.txt", header = TRUE, sep = ";", na.strings = "?")
```

Let's take a look at the data
```{r}
class(data)
```
```{r}
head(data)
```
We are only interested in the data from 2007-02-01 to 2007-02-02. Let's subset the data
```{r}
# Convert Date and Time variables to Date/Time classes in R using the strptime() and as.Date() functions.
data$Date <- as.Date(data$Date, format = "%d/%m/%Y")
data$Time <- strptime(data$Time, format = "%H:%M:%S")

# Subset the data
data_sub <- subset(data, Date %in% as.Date(c("2007-02-01", "2007-02-02")))

# Combine the Date and Time variables to create a single Date/Time variable
data_sub$DateTime <- as.POSIXct(paste(data_sub$Date, format(data_sub$Time, "%H:%M:%S")), format = "%Y-%m-%d %H:%M:%S")
```

Let's take a look at the subsetted data
```{r}
print(class(data_sub))
head(data_sub)
```


## Plots

### Plot 1
Our goal is to examine how household energy usage varies over a 2-day period in February, 2007.
```{r}
# Plot 1
source("plot1.R")
plot1(data_sub)

# Save the plot
dev.copy(png, "plot1.png", width = 480, height = 480)
```

### Plot 2
Our goal is to examine how `Global Active Power` varies over a 2-day period in February, 2007 on Thursday, Friday, and Saturday.
```{r}
# Plot 2
# Plot a line graph of Global Active Power over the 2-day period
# Combine the Date and Time variables to create a single Date/Time variable
source("plot2.R")
plot2(data_sub)
dev.copy(png, "plot2.png", width = 480, height = 480)
```

### Plot 3
Our goal is show Energy Sub Metering over the 2-day period in February, 2007 factored by sub metering type.
```{r}
# Plot 3
source("plot3.R")
plot3(data_sub)
dev.copy(png, "plot3.png", width = 480, height = 480)
```

### Plot 4
The goal is to plot a 2 x 2 grid of plots showing the following:
- Plot 2 in the upper left corner
- A new plot containing `Voltage` vs `DateTime` in the upper right corner
- Plot 3 in the lower left corner
- A new plot containing `Global Reactive Power` vs `DateTime` in the lower right corner

```{r}
# Plot 4
source("plot4.R")
plot4(data_sub)
dev.copy(png, "plot4.png", width = 480, height = 480)
```
